---
title: "sdcmicro-exercise"
author: "YOUR NAMES HERE"
date: "2023-05-24"
output: html_document
---

# Whale Entanglement sdcMicro Exercise

Your team acquired a dataset\* [whale-sdc.csv](whale-sdc.csv) from researchers working with whale entanglement data on the West Coast. The dataset contains both direct and indirect identifiers. Your task is to assess the risk of re-identification of the fisheries associated with the cases before considering public release. Then, you should test one technique and apply k-anonymization to help lower the disclosure risk as well as compute the information loss.

Please complete this exercise in pairs or groups of three. Each group should download the dataset and complete the rmd file, including the code and answering the questions. Remember to include your names in the YAML.

*\*This dataset was purposefully adapted exclusively for instruction use.*

#### *Setup*

#### Package & Data

```{r}
library(here)
library(tidyverse)
library(sdcMicro)

wd <- here("week9")
```

#### Inspect the Dataset

```{r}
#Read the CSV dataset into a data frame
whale_data <- read.csv(here(wd, "whale-sdc.csv"))
whale_df <- whale_data # keeping raw data as-is

#Show the list of variable names
head(whale_df)

#Summary of the data frame
str(whale_df)

```

#### Q1. How many direct identifiers are present in this dataset? What are they?

A: 
- fishery_license
- lat + long
- case_id (to keep internally)

#### **Q2. What attributes would you consider quasi-identifiers? Why?**

A: 
- county (there can be multiple fisheries in  a county, location), 
- state 
- (NOT lat, long) (multiple fisheries can fish in the same area, whales can be found in multiple locations that overlap w multiple fishery areas, but some feel where they fish is private data)
- origin (could be multiple fisheries with the same origin but helps to narrow it down a lot)

#### **Q3. What types of variables are they? Define them. (numeric, integer, factor or string)**

fishery_license: numeric/double (should be factor but we're getting rid of it so no need to convert)
county: character (should be factor?)
lat & long: numeric 
year & month: could be factors, but we're setting to numeric

also want  type, state, condition, origin, gear, fine, infraction type to be factors
same with inj_level

Make sure to have them set correctly.

```{r}
# Convert variables to factors
whale_data <- varToFactor(obj=whale_df, var=c("county", "type", "state", "condition", "origin", "gear", "fine", "infraction_type", "inj_level"))

whale_data <- varToNumeric(obj = whale_data, var = c("lat", "long", "year", "month"))

```

#### ***4 Considering your answers to questions 1, 2 and 3 create a SDC problem.***

#### **Q4.1 What is the risk of re-identification for this dataset?**

```{r}
#Select variables for creating the sdcMicro object
#Select categorical key variables - aka quasi-identifiers

sdc_initial <- createSdcObj(dat = whale_data,
                           keyVars = c("county", "state", "origin"),
                           numVars = c("year", "month"),
                           weightVar = NULL,
                           hhId = NULL,
                           strataVar = NULL,
                           pramVars = NULL,
                           # keeping case_id for now, would want to drop at the end
                           excludeVars = c("fishery_license", "lat", "long"),
                           seed = 0,
                           randomizeRecords = FALSE,
                           alpha = c(1))
```


```{r}
# Summary of object
sdc_initial
summary(sdc_initial)


sdc_initial@risk$global
sdc_initial@risk$global$risk

paste0("The risk of identifying a fishery based on the current data is ", round(sdc_initial@risk$global$risk_pct, 3), "%.")

head(sdc_initial@risk$individual)
```

#### Q4.2 To what extent does this dataset violate k-anonymity?

```{r}
sdc_initial
print(sdc_initial, 'kAnon')

```

#### *5. Consider techniques that could reduce the risk of re-identification.*

#### Q5.1 Apply one non-perturbative method to a variable of your choice. How effective was it in lowering the disclosure risk?

```{r}
table(sdc_initial@manipKeyVars$county)

table(sdc_initial@manipKeyVars$origin)

# dropping those 8 rows of recreational and tribal might be better, or dropping the column (not as recommended)
sdc_initial <- groupAndRename(obj= sdc_initial, var= c("origin"), before=c("recreational", "tribal"),
                              after=c("non-commercial")) 

# check to see if binning worked
table(sdc_initial@manipKeyVars$origin)

# paste0("The risk of identifying a fishery based on the updated data is ", round(sdc_initial@risk$global$risk_pct, 3), "%.")
sdc_initial
```

#### Q5.2 Apply ( k-3) anonymization to this dataset.

```{r}
#Local suppression to obtain k-anonymity
sdc_initial <- kAnon(sdc_initial, k=c(3))

#sdc_initial

#Global risk
sdc_initial@risk$global$risk_pct # down to 5.07% !!!

#K-anon

print(sdc_initial, 'kAnon')
```

#### Q6. Compute the information loss for the de-identified version of the dataset. 

```{r}
print(sdc_initial, 'ls')

```
